Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-07-17 manuscript/manuscript.tex
updates ms with new structure

Forecasting future states of populations has taken on new urgency as the
rate of climate change increases. Traditional plant population models
have limited utility in this regard because they are based on detailed
demographic data from small, localized plots. These models are difficult
to scale up to spatial scales relevant to land managers that require
such forecasts to make decisions. To overcome the data limitations of
traditional population models, some have proposed population models
based on population level, rather than individual level, data that is
much easier to collect over broad spatial scales. Using such models
violates a central assumption of ecology: individuals respond to
weather, not populations to climate. Here, we test whether this
assumption is important when forecasting climate change impacts on four
perennial grass species in a semi-arid Montana grassland. We
parameterized two population models, one based on inidividual level data
with three vital rates and one on an aggregated version of the same data
(percent cover), and compared their accuracy, precision, and sensitivity
to climate. The individual level model was more accurate and precise
than the aggregate level model when predicting out of sample
observations. The aggregate level model made countervailing forecasts to
1\% climate changes when compared to the individual level model,
indicating the aggregate level model produces incorrect forecasts. When
comparing climate effects from both models, the aggregate level model
tends to ``miss'' important climate effects from at least one vital rate
for each species. It appears there is no short cut to forecasting
climate change impacts on plant populations --- detailed demographic
data is essential. But, forecasts were very uncertain, so we advocate
for a focus on new methods to collect demographic data more efficiently
across environmental gradients in space and time.

@@

Population models are important tools for predicting the impacts of
environmental change on species persistence and abundance. But
reconciling the scales at which population models are parameterized and
the scales at which environmental changes play out remains a challenge
(Clark et al. 2010, 2012, Freckleton et al. 2011, Queenborough et al.
2011). The major hurdle is that most population models are built using
data from a single study site because collecting those data, which
involves tracking the fates of individuals plants, is so difficult. The
resulting models cannot be applied to the landscape and regional scales
relevant to decision-makers without information about how the fitted
parameters respond to spatial variation in biotic and abiotic drivers
(Sæther et al. 2007). The temporal coverage of demographic datasets is
no better. {[}I think the temporal extent is a separate issue, might
muddy the waters to bring it up here\ldots{}{]} For example, the most
common study duration in the COMPADRE matrix population model database
is 4 years and only a few studies exceed 10 years (Salguero-Gómez et al.
2015). The limited spatio-temporal extent of individual-level
demographic datasets constrains our ability to use population models to
address applied questions about the consequences of climate change.

The inability of many population models to address landscape-scale
problems may explain why land managers and conservation planners have
embraced species distribution models (SDMs) (see Guisan and Thuiller
2005 for a review). SDMs typically rely on easy-to-collect
presence/absence data (but see Clark et al. 2014 for new methods) and
remotely-sensed environmental covariates that allow researchers to model
large spatial extents (e.g., Maiorano et al. 2013). Thus, it is
relatively straightforward to parameterize and project SDMs over
landscapes and regions, the scales at which many land-use decisions are
made. However, the limitations of SDMs are well known (Elith and
Leathwick 2009). Ideally, researchers would provide managers with
landscape-scale population models, combing the extent of SDMs with
information about dynamics and species abundances.

@@

Our main goal in this paper is to see if models based on aggregated data
are as sensitive to climate covariates as models based on individual
level data. So, with our fitted and validated models in hand, we ran
simulations for each model type (IPM and QBM) under four climate
perturbation scenarios: (1) observed climate, (2) precipitation
increased by 1\%, (3) temperature increased by 1\%, and (4)
precipitation and temperature increased by 1\%. We ran the simulations
for 2,500 time steps, enough to estimate equilibrium cover after
discarding an initial 500 time steps as burn-in. Each simulation was run
under two parameter scenarios: (1) using mean parameter estimates and
(2) using randomly drawn parameters from the MCMC chain. We use (1) to
detect the overall sensitivity of equilibrium cover to climate, and we
use (2) to show the impact of model uncertainty on forecast precision.

@@

Our \emph{a priori} expectation is that the IPM will produce more
accurate and precise forecasts. This could be due to differences in
sample sizes leading to larger parameter uncertainty for the QBM, or due
to the QBM climate effects being weakly associated with one or more
vital rate climate effects at the individual level. To assess the impact
of sample size on QBM parameter uncertainty we refit the QBM statistical
model (Eqs. 8-9) after removing sets of 2, 5, 10, and 15 quadrats. We
fit 10 models at each level of quadrat removal (2, 5, 10, 15 quadrats),
removing a different randomly selected set of quadrats for each fit. We
calculated the standard deviation of climate main effects (pptlag, ppt1,
ppt2, TmeanSpr1, and TmeanSpr2) for each model and averaged those over
replicates within each set of quadrat removals. This allowed us to
regress sample size against parameter uncertainty.

@@

Equilibrium cover from both models was sensitive to climate (Fig. 3a-d).
The IPM projected percent changes in equilibrium cover from -3 to 8\%
for \emph{B. gracilis}, -4 to 3\% for \emph{H. comata}, -15 to 9\% for
\emph{P. smithii}, and -17 to 53\% for \emph{P. secunda}. The QBM
projected opposite and greater percent changes in equilibrium cover for
\emph{B. gracilis} (-63 to 30\%) and \emph{H. comata} (-50 to -18\%;
Fig. 3a-b). For \emph{P. smithii}, the QBM projected opposite changes in
equilibrium cover than the IPM, but of similar magnitude (-5 to 6\%;
Fig. 3c). \emph{P. secunda} was the only species that the IPM and QBM
made projections of the same sign and somewhat similar magnitude (-20 to
14\%; Fig. 3d).

The response of a population to climate change is a result of the
aggregate effects of climate on individual vital rates. Since the IPM
approach relies on vital rate regressions, we were able to quantify the
sensitivity of each vital rate in isolation and in pairwise
combinations. Across all species, climate covariates can have opposing
effects on different vital rates (Fig. 3e-h).

Growth was the most sensitive vital rate for all species, showing a
negative response to increased precipitation, and stronger positive
response to increased temperature, and a mostly positive response when
both climate factors are increased (Fig. 3e-h). \emph{B. gracilis}
survival rates were sensitive to temperature, resulting in an increase
in plant cover under increased temperature (Fig. 3e). In isolation,
recruitment and survival were insensitive to climate factors for
\emph{H. comata} (Fig. 3f). Survival and recruitment of \emph{P.
smithii} were both sensitive, negatively, to temperature and
precipitation (Fig. 3g). \emph{P. secunda} equilibrium cover was
sensitive to the climate effects on survival and recruitment, showing a
negative effect on both vital rates for increased precipition, but a
strong positive effect on survival with increased temperature (Fig. 3h).
The climate impact of recruitment on equilibrium cover was negative for
precipitation and temperature increases (Fig. 3h). At least two of three
vital rates were sensitive to climate for each species (Fig. 3).

Forecasts based on 1\% climate changes were extremely uncertain when we
considered model error and parameter uncertainty (Fig. 4). As expected
based on model validation (Table 1), QBM projections were more uncertain
than IPM projections for all species except \emph{P. smithiii} (Fig. 4).
IPM forecasts for \emph{P. smithiii} were very uncertain due to
overcompensating density dependence and nonlinearities recruitment that
get amplified by parameter uncertainty (Appendx C). Neither model was
capable of making forecasts of proportional cover change distinguishable
from 0 (i.e., no change in population state) when we included model
error and parameter uncertainty.

@@

@@

\subsubsection{The IPM and QBM produced inconsistent
forecasts}\label{the-ipm-and-qbm-produced-inconsistent-forecasts}

Using individual-level and population-level forms of the same dataset,
we were able to directly compare a traditional demographic modeling
approach to a population model based on percent cover data. Our
quad-based model (QBM) is based on percent cover data and so is in the
spirit of density-structured models. In terms of each model's
forecasting ability, the IPM outperformed the QBM (Table 1). This is
unsurprising since we expected to lose some precision at the aggregated
level. However, the underwhelming performance of the QBM could call into
the question forecasts that differ from the IPM.

Indeed, when we perturbed climate factors the QBM made forecasts
completely contradictory to those of the IPM for three of our four study
species (Fig. 3a-d). In a perfect world, the QBM would have made
forecasts of at least the same direction as the IPM. If that had been
the case we could conclude that aggregate level models could prove
useful for forecasting climate change impacts on populations.
Unfortunately, this was not the case.

Given the superior ability of the IPM to predict out of sample
observations (Table 1), we have no choice but to conclude it is the
superior model. Following that logic, we can only assume that, at least
contingent on the data in hand, the IPM is producing the correct
forecasts to climate perturbations. The QBM failed to match IPM
forecasts, implying that detailed demographic data may be necessary to
accurately detect climate signals that are utlimately important at the
population level. This result further confirms related work on the
importance of individual variability on population level responses to
exogenous drivers (Clark et al. 2011a, 2011b, 2012, Galván et al. 2014).

\subsubsection{The role of vital rate climate
dependence}\label{the-role-of-vital-rate-climate-dependence}

We can think of two reasons why the IPM outperformed the QBM. First, the
quadrat level data has a much reduced sample size compared to the
individual level data. In an ideal world we would have compared the IPM
and QBM using data collected over the same amount of person hours, not
just the same number of quadrats. Then the sample size of the quadrat
level data would be much greater and carry more statistical power. To
address this limitation in our work we fit the QBM statistical model
(Eq. 8-9) with different numbers of quadrats to see the effect of sample
size on the precision of climate effect estimates. It appears that
including additional quadrats leads to rapidly diminishing returns in
terms of parameter precision (Fig. 5). Thus, while sample size surely
plays some role, we do not think it is the main driver of the difference
between the IPM and QBM.

The second reason the IPM could have outperformed the QBM is that the
population level model is in fact missing important climate effects that
act on individual vital rates, rather than population growth. Our
intuition was that species with strong climate-dependence on vital rates
not well resolved at the aggregate level would result in different
forecasts from the two models. For example, survival is very size
dependent: smaller individuals have a higher probability of death (Chu
and Adler 2014). At the same time, a single small individual contributes
relatively little to percent cover estimates at the plot scale. So, if
survival of individuals was positively impacted by temperature
increases, for example, we would expect to detect this signal in the
individual level data but not in the aggregate level data. To see if
this is the case we can regress climate effects from each vital rate
statistical model at the individual level against the same climate
effects from the QBM statistical model (Fig. 6).

In general, the QBM climate effects are most correlated with climate
effects from the growth regression at the individual level (Fig. 6). In
no case does the QBM statistical model have strong correlations across
all three vital rates (Fig. 6). Thus, for each species the QBM is
``missing'' climate signals associated with at least one vital rate.
This has large impacts on predictions of long term population dynamics,
as seen in our equilibrium simulations (Fig. 3a-d). The inability of the
QBM to separate the sometimes countervailing effects of climate on
survival, growth, and recruitment (Fig. 3e-h) results in inaccurate
(Fig. 3a-d) and unprecise (Fig. 4) forecasts. The QBM statistical model
struggles to explain variation due to climate variables because they can
have positive and negative impacts on different vital rates. When this
is the case, as it is for all our species to varying degrees (Fig.
3e-h), statistical models of aggregated population responses will fail.

These results lead us to conclude that detailed demographic data is
necessary to forecast climate change impacts on plant populations. This
is unwelcome news since this data is difficult to collect and the models
built on such data are of little use to land managers that make
decisions at scales beyond that of traditional population models
(Queenborough et al. 2011). While density-structured approaches may fail
when climate covariates are considered, there are other alternatives.
For example, Clark et al. (2011a) use Forest Inventory and Analysis
(FIA) data to parameterize a population model with multiple vital rates
and climate dependence. Another example are distributed efforts like
PlantPopNet (\url{http://plantago.plantpopnet.com}) that will allow
researchers to estimate variation around climate responses for
widespread species by taking advantage of spatial variation in climate
(e.g. Doak and Morris 2010). Lastly, we foresee new approaches on the
horizon that leverage photo/video of plots and advanced object
recognition algorithms (e.g. Liu et al. 2014) to streamline plant
mapping and digitizing efforts.

\subsubsection{Forecasting the future, and the future of
forecasting}\label{forecasting-the-future-and-the-future-of-forecasting}

Our goal was not to make any explicit forecast for the future state of
these populations based on predicted climate change. But our results
highlight the state of affairs in ecology when it comes to forecasting
the impacts of climate change. The analysis we conducted here could be
considered, with some exceptions of course, at the forefront of
ecological forecasting in terms of the statistical approach employed
(hierarchical Bayesian), the type of population model we used
(stochastic IPM with parameter uncertainty), and the amount of high
quality data we had at our disposal (14 years of individual level data).
Yet, model predictions proved so uncertain that any forecast, when
bounded with model and parameter uncertainty, would be at best not
useful and at worst meaningless. For all species, the 90\% quantiles of
predicted changes in populatoin size overlapped zero; we cannot even
predict whether a change in precipitation or temperature will cause
populations to increase or decrease. How might we improve on this state
of affairs?

First, forecasts could be improved by matching the spatial scale of
predictor variables with the spatial scale of observations. One of the
major limitations of the models we fit here is that the climate data are
at a much larger scale than the individual level observations of plant
size. Climate covariates only vary by year, with no spatial variability
within years. Thus, even if we fit models to individual level data, we
are missing the key interaction point between weather and individual
plants (Clark et al. 2011b) because all observations share the same
climate covariates. Demographic studies should be designed with at least
plot level measurements of climate related variables (e.g., soil
moisture).

@@

Third, ecologists as a community need to get serious about reporting
uncertainty. There is a strong culture around explicitly considering
model uncertainty, but parameter uncertainty is often ignored. In some
cases this is because the easiest statistical methods do no make
propogating parameter uncertainty a straighforward task. Even Bayesian
approaches that allow integration of model fitting and forecasting
(Hobbs and Hooten 2015) are not simple when using modeling approaches
like integral projection models that separate the model fitting and
simulation stages (Rees and Ellner 2009). However, as we have done here,
it is still possible to include parameter uncertainty by drawing
parameter values from MCMC iterations, taking care to draw all
parameters from the same chain and iteration to account for their
correlations. Only by being honest about our forecasts can we begin to
produce better ones.

@@

This work is not a critique of density-structured population models. In
some cases and for certain species, population models based on
aggregated data may prove useful and unbiased. However, our work here is
the first comparison, to our knowledge, of population models based on
individual and aggregated forms of the same data in a climate change
context. Our results confirm theoretical arguments (Clark et al. 2011b)
and empirical evidence (Clark et al. 2011a, 2012) that individual
responses are critical to predicting species' responses to climate
change. Thus, forecasts from aggregate level models should be viewed
with caution and should never be unaccompianed by uncertainty. Given the
importance of demographic data and its current difficulty to collect, we
advocate for research on modern methods to collect demographic data more
efficiently across environmental gradients in space and time.

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_3.pdf}
\caption{Proportional change in species' mean cover caused by a 1\%
increase in observed precipitation (+ppt), temperature (+temp), or both
(+ppt\&temp) as predicted by the individual-based IPM and the
aggregate-based QBM using mean parameter values. Top panels show the
mean predicted proportional change in cover. Lower panels show the
sensitivity of equilibrium cover simulated from the IPM to each climate
scenario applied to individual and combined vital rates. For example,
the points associated with G show the median cover from IPM simulations
where a climate perturbation is applied only to the growth regression
climate covariates. These simulations also use mean parameter values for
clarity.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_4.pdf}
\caption{Equilibrium cover and 90\% quantiles around the mean prediction
when model error and parameter uncertainty are propogated through the
simulation phase. Climate simulations are as in Figure 3.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_5.pdf}
\caption{Effect of quadrat sample size on the precision (standard
deviation) of main climate effect estimates in the QBM. Increasing the
number of quadrats results in diminishing returns in terms of parameter
certainty. Light dashed lines show individual climate effects at five
quadrat sample sizes. Thick dark lines are inverse gaussian fits showing
the mean effect of increasing quadrat sample size on parameter
precision.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_6.pdf}
\caption{Correlations (r) between QBM and IPM estimates of climate
effects. We ignore sizeXclimate interactions since these are not
directly comparable across model types. The QBM does not have multiple
vital rates, so its values are repeated across panels within each
species. Across top panels, `growth' = growth regression, `rec' =
recruitment regression, `surv' = survival regression.}
\end{figure}

@@

@@

@@

@@

Salguero-Gómez, R., O. R. Jones, C. R. Archer, Y. M. Buckley, J.
Che-Castaldo, H. Caswell, D. Hodgson, A. Scheuerlein, D. A. Conde, E.
Brinks, H. de Buhr, C. Farack, F. Gottschalk, A. Hartmann, A. Henning,
G. Hoppe, G. Römer, J. Runge, T. Ruoff, J. Wille, S. Zeh, R. Davison, D.
Vieregg, A. Baudisch, R. Altwegg, F. Colchero, M. Dong, H. de Kroon,
J.-D. Lebreton, C. J. E. Metcalf, M. M. Neel, I. M. Parker, T. Takada,
T. Valverde, L. A. Vélez-Espino, G. M. Wardle, M. Franco, and J. W.
Vaupel. 2015. The compadrePlant Matrix Database: an open online
repository for plant demography. Journal of Ecology 103:202--218.

>>>>

Forecasting future states of populations has taken on new urgency as the
rate of climate change increases. Traditional plant population models
have limited utility in this regard because they are based on detailed
demographic data from small, localized plots. These models are difficult
to scale up to spatial scales relevant to land managers that require
such forecasts to make decisions. To overcome the data limitations of
traditional population models, some have proposed population models
based on population level, rather than individual level, data that is
much easier to collect over broad spatial scales. Using such models
violates a central assumption of ecology: individuals respond to
weather, not populations to climate. Here, we test whether this
assumption is important when forecasting climate change impacts on four
perennial grass species in a semi-arid Montana grassland. We
parameterized two population models, one based on inidividual-level data
with three vital rates and one on population-level data from the same
dataset (percent cover of quadrats), and compared their accuracy,
precision, and sensitivity to climate. The individual level model was
more accurate and precise than the aggregate level model when predicting
out-of-sample observations. When comparing climate effects from both
models, the population-level model tends to ``miss'' important climate
effects from at least one vital rate for each species. We also find that
increasing the sample size at the population-level would not necessarily
reduce forecast certainty; meaning the only way to reduce uncertainty is
to capture unique climate dependence on individual vital rates. It
appears there is no short cut to forecasting climate change impacts on
plant populations --- detailed demographic data is essential. However,
forecasts from our individual-level model were very uncertain, so we
advocate for a focus on new methods to collect demographic data more
efficiently across environmental gradients in space and time.

@@

Population models are important tools for predicting the impacts of
environmental change on species persistence and abundance. But
reconciling the scales at which population models are parameterized and
the scales at which environmental changes play out remains a challenge
(Clark et al. 2010, 2012, Freckleton et al. 2011, Queenborough et al.
2011). The major hurdle is that most population models are built using
data from a single study site because collecting those data, which
involves tracking the fates of individuals plants, is so difficult. The
resulting models cannot be applied to the landscape and regional scales
relevant to decision-makers without information about how the fitted
parameters respond to spatial variation in biotic and abiotic drivers
(Sæther et al. 2007). The limited spatial extent of individual-level
demographic datasets constrains our ability to use population models to
address applied questions about the consequences of climate change.

The inability of many population models to address landscape-scale
problems may explain why land managers and conservation planners have
embraced species distribution models (SDMs) (see Guisan and Thuiller
2005 for a review). SDMs typically rely on easy-to-collect
presence/absence data (but see Clark et al. 2014 for new methods) and
remotely-sensed environmental covariates that allow researchers to model
large spatial extents (e.g., Maiorano et al. 2013). Thus, it is
relatively straightforward to parameterize and project SDMs over
landscapes and regions, the scales at which many land-use decisions are
made. However, the limitations of SDMs are well known (Pearson and
Dawson 2003, Elith and Leathwick 2009, Araújo and Peterson 2012).
Ideally, researchers would provide managers with landscape-scale
population models, combing the extent of SDMs with information about
dynamics and species abundances.

@@

Our main goal in this paper is to see if models based on aggregated data
are as sensitive to climate covariates as models based on individual
level data. So, with our fitted and validated models in hand, we ran
simulations for each model type (IPM and QBM) under four climate
perturbation scenarios: (1) observed climate, (2) precipitation
increased by 1\%, (3) temperature increased by 1\%, and (4)
precipitation and temperature increased by 1\%. We ran the simulations
for 2,500 time steps, enough to estimate equilibrium cover after
discarding an initial 500 time steps as burn-in. Each simulation was run
under two parameter scenarios: (1) using mean parameter estimates and
(2) using randomly drawn parameters from the MCMC chain. We use (1) to
detect the overall sensitivity of equilibrium cover to climate, and we
use (2) to show the impact of model and parameter uncertainty on
forecast precision.

@@

Our expectation is that the IPM will produce more accurate and precise
forecasts. This could be due to differences in sample sizes leading to
larger parameter uncertainty for the QBM, or due to the QBM climate
effects being weakly associated with one or more vital rate climate
effects at the individual level. To assess the impact of sample size on
QBM parameter uncertainty we refit the QBM statistical model (Eqs. 8-9)
after removing sets of 2, 5, 10, and 15 quadrats. We fit 10 models at
each level of quadrat removal (2, 5, 10, 15 quadrats), removing a
different randomly selected set of quadrats for each fit. We calculated
the standard deviation of climate main effects (pptlag, ppt1, ppt2,
TmeanSpr1, and TmeanSpr2) for each model and averaged those over
replicates within each set of quadrat removals. This allowed us to
regress sample size against parameter uncertainty.

@@

The response of a population to climate change is a result of the
aggregate effects of climate on individual vital rates. Since the IPM
approach relies on vital rate regressions, we were able to quantify the
sensitivity of each vital rate in isolation and in pairwise
combinations. Across all species, climate covariates can have opposing
effects on different vital rates (Fig. 3). Growth was the most sensitive
vital rate for all species, showing a negative response to increased
precipitation, and stronger positive response to increased temperature,
and a mostly positive response when both climate factors are increased
(Fig. 3). \emph{B. gracilis} survival rates were sensitive to
temperature, resulting in an increase in plant cover under increased
temperature (Fig. 3a). In isolation, recruitment and survival were
insensitive to climate factors for \emph{H. comata} (Fig. 3b). Survival
and recruitment of \emph{P. smithii} were both sensitive, negatively, to
temperature and precipitation (Fig. 3c). \emph{P. secunda} equilibrium
cover was sensitive to the climate effects on survival and recruitment,
showing a negative effect on both vital rates for increased precipition,
but a strong positive effect on survival with increased temperature
(Fig. 3d). The climate impact of recruitment on equilibrium cover was
negative for precipitation and temperature increases (Fig. 3d). At least
two of three vital rates were sensitive to climate for each species
(Fig. 3).

@@

\subsubsection{Model forecasts}\label{model-forecasts}

Forecasts based on 1\% climate changes were extremely uncertain when we
considered model error and parameter uncertainty (Fig. 6). As expected
based on model validation (Table 1), QBM projections were more uncertain
than IPM projections for all species except \emph{P. smithiii} (Fig. 6).
IPM forecasts for \emph{P. smithiii} were very uncertain due to
overcompensating density dependence and nonlinearities recruitment that
get amplified by parameter uncertainty (Appendx C). Neither model was
capable of making forecasts of proportional cover change distinguishable
from 0 (i.e., no change in population state) when we included model
error and parameter uncertainty.

@@

Using individual-level and population-level forms of the same dataset,
we were able to directly compare a traditional demographic modeling
approach to a population model based on percent cover data. Our
quad-based model (QBM) is based on percent cover data and so is in the
spirit of density-structured models. In terms of each model's
forecasting ability, the IPM outperformed the QBM: the IPM was more
accurate and precise than the QBM (Table 1). As we suggest in the
introduction, this could be due to differences in sample size or the
effect of averaging over climate responses unique to individual-level
vital rates.

We found no evidence that increasing sample size at the quadrat level
would lead to higher precision of climate coefficient estimates (Fig.
4). We did, however, find evidence that the QBM statistical model failed
to identify climate dependence for some vital rates (Fig. 5). For no
species were climate effects from the QBM strongly correlated with all
three vital rates (Fig. 5). Freckleton et al. (2011) acknowledge that
averaging over complex stage dependence will lead to poorly specified
models. This is analagous to our situation, but instead of averaging
over complex life histories, we are averaging over complex climate
dependence.

Thus, for each species the QBM is ``missing'' climate signals associated
with at least one vital rate. This leads to inaccurate and imprecise
forecasts because the QBM statistical model struggles to explain
variation due to climate variables that have positive and negative
impacts on different vital rates. When this is the case, as it is for
all our species to varying degrees (Fig. 3), statistical models of
population-level data will fail. This result further confirms related
work on the importance of individual-level data to forecast population
responses to exogenous drivers (Clark et al. 2011a, 2011b, 2012, Galván
et al. 2014).

These results lead us to conclude that detailed demographic data is
necessary to forecast climate change impacts on plant populations when
vital rates have unique climate responses. This is unwelcome news since
this data is difficult to collect and the models built on such data are
of little use to land managers that make decisions at scales beyond that
of traditional population models (Queenborough et al. 2011). While
density-structured approaches may fail when climate covariates are
considered, there are other alternatives. For example, Clark et al.
(2011a) use Forest Inventory and Analysis (FIA) data to parameterize a
population model with multiple vital rates and climate dependence.
Another example are distributed efforts like PlantPopNet
(\url{http://plantago.plantpopnet.com}) that will allow researchers to
estimate variation around climate responses for widespread species by
taking advantage of spatial variation in climate (e.g. Doak and Morris
2010). Lastly, we foresee new approaches on the horizon that leverage
photo/video of plots and advanced object recognition algorithms (e.g.
Liu et al. 2014) to streamline plant mapping and digitizing efforts.

What does all this mean for model forecasts? The answer is more nuanced
than one might expect. The typical approach in ecology is to use point
estimates of model parameters to project populations forward according
to the specified model; usually allowing for some variability around the
determinstic process (e.g., Battin et al. (2007); Jenouvrier et al.
(2009); Adler et al. (2012)). If we follow tradition and calculate the
mean response to climate change over many iterations of the model with
only model error and temporal variation included, the IPM and the QBM
produce opposing forecasts for three of four species (Fig. D1). However,
if we introduce parameter uncertainty, the forecasts are actually
indistinguishable (Fig. 6). IPM forecasts are generally more precise, as
expected from model validation (Table 1). The real story is not that the
QBM produces ``wrong'' forecasts, but that it produces highly uncertain
forecasts. Our sample size analysis reveals that this uncertainty is not
easily overcome by including more data (Fig. 4). The only way to reduce
uncertainty is to model these populations based on individual-level data
where climate dependence can be properly identified (Fig. 5). Yet, even
the IPM forecasts were highly variable (Fig. 6).

Our goal was not to make any explicit forecast for the future state of
these populations based on predicted climate change. But our results
highlight the state of affairs in ecology when it comes to forecasting
the impacts of climate change. The analysis we conducted here could be
considered, with some exceptions of course, at the forefront of
ecological forecasting in terms of the statistical approach employed
(hierarchical Bayesian), the type of population model we used
(stochastic IPM with parameter uncertainty), and the amount of high
quality data we had at our disposal (14 years of individual level data).
Yet, model predictions proved so uncertain that any forecast, when
bounded with model and parameter uncertainty, would be at best not
useful and at worst meaningless. For all species, the 90\% quantiles of
predicted changes in population size overlapped zero; we cannot even
predict whether a change in precipitation or temperature will cause
populations to increase or decrease. How might we improve on this state
of affairs?

First, forecasts could be improved by matching the spatial scale of
predictor variables with the spatial scale of observations. One of the
major limitations of the models we fit here is that the climate data are
at a much larger scale than the individual level observations of plant
size. Climate covariates only vary by year, with no spatial variability
within years. Thus, even if we fit models to individual-level data, we
are missing the key interaction point between weather and individual
plants (Clark et al. 2011b) because all observations share the same
climate covariates. Demographic studies should be designed with at least
plot level measurements of climate related variables (e.g., soil
moisture).

@@

Third, ecologists as a community need to get serious about reporting
uncertainty. There is a strong culture around explicitly considering
model uncertainty, but parameter uncertainty is often ignored. In some
cases this is because the easiest statistical methods do no make
propogating parameter uncertainty a straighforward task. Even Bayesian
approaches that allow integration of model fitting and forecasting
(Hobbs and Hooten 2015) are not simple when using modeling approaches
like integral projection models that separate the model fitting and
simulation stages (Rees and Ellner 2009). However, as we have done here,
it is still possible to include parameter uncertainty by drawing
parameter values from MCMC iterations, taking care to draw all
parameters from the same chain and iteration to account for their
correlations. Only by being honest about our forecasts can we begin to
produce better ones, and forecasts reported without parameter error are
disingenuous. In many cases, ignoring parameter error is justifiable
when investigating basic demographic mechanisms, but it is never
justifiable if forecasts are being made.

@@

This work is not a critique of density-structured population models. In
some cases and for certain species where climate dependence is weak or
concentrated within a single vital rate, population models based on
aggregated data may prove useful and unbiased. However, our work here is
the first comparison, to our knowledge, of population models based on
individual and aggregated forms of the same data in a climate change
context. Our results confirm theoretical arguments (Clark et al. 2011b)
and empirical evidence (Clark et al. 2011a, 2012) that individual
responses are critical to predicting species' responses to climate
change. Thus, forecasts from population-level models should be viewed
with caution and should never be unaccompianed by uncertainty. Given the
importance of demographic data and its current difficulty to collect, we
advocate for research on modern methods to collect demographic data more
efficiently across environmental gradients in space and time.

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_3.pdf}
\caption{Sensitivity of equilibrium cover simulated from the IPM to each
climate scenario applied to individual and combined vital rates. For
example, the points associated with G show the median cover from IPM
simulations where a climate perturbation is applied only to the growth
regression climate covariates. These simulations use mean parameter
values for clarity.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_4.pdf}
\caption{Effect of quadrat sample size on the precision (standard
deviation) of main climate effect estimates in the QBM. Increasing the
number of quadrats results in diminishing returns in terms of parameter
certainty. Light dashed lines show individual climate effects at five
quadrat sample sizes. Thick dark lines are inverse gaussian fits showing
the mean effect of increasing quadrat sample size on parameter
precision.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_5.pdf}
\caption{Correlations (r) between QBM and IPM estimates of climate
effects. We ignore sizeXclimate interactions since these are not
directly comparable across model types. The QBM does not have multiple
vital rates, so its values are repeated across panels within each
species. Across top panels, `growth' = growth regression, `rec' =
recruitment regression, `surv' = survival regression.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_6.pdf}
\caption{Equilibrium cover and 90\% quantiles around the mean prediction
when model error and parameter uncertainty are propogated through the
simulation phase. Climate simulations are as in Figure 3.}
\end{figure}

@@

Araújo, M. B., and A. T. Peterson. 2012. Uses and misuses of bioclimatic
envelope modeling. Ecology 93:1527--1539.

Battin, J., M. W. Wiley, M. H. Ruckelshaus, R. N. Palmer, E. Korb, K. K.
Bartz, and H. Imaki. 2007. Projected impacts of climate change on salmon
habitat restoration. Proceedings of the National Academy of Sciences of
the United States of America 104:6720--6725.

@@

Jenouvrier, S., H. Caswell, C. Barbraud, M. Holland, J. Stroeve, and H.
Weimerskirch. 2009. Demographic models and IPCC climate projections
predict the decline of an emperor penguin population. Proceedings of the
National Academy of Sciences of the United States of America
106:1844--1847.

@@

Pearson, R. G., and T. P. Dawson. 2003. Predicting the impacts of
climate change on the distribution of species: Are bioclimate envelope
models useful? Global Ecology and Biogeography 12:361--371.

@@
